---
title: "Central Limit Theorem Simulations"
author: "Alim Ray"
date: "9/26/2015"
output:
  html_document: default
---

#Overview
This report explores the Cental Limit Theorem by running simulations using the exponential distribution.  The sample means of the simulations
are plotted and shown to fix the normal distribution.


#Simulation

Below is code that runs the exponential distribution 1,000 times.  Cumulative means for the 1,00 simulations are plotted and shown to converge
to the theorectical mean 1/\lambda where \lambda = 0.2.


```{r}
library(ggplot2)

simulations <- function( nosim = 1000 ) {
  n <- 40
	lambda <- 0.2
	mns <- c()
  vrs <- c()
	for (i in 1 : nosim) {
		sim <- rexp(n, lambda)
		mns <- c( mns, mean(sim) )
		vrs <- c( vrs, var(sim) )
  }
  list( mns, vrs )
}

simulationGraphs <- function(nosim = 1000) {
  simulationValues <- simulations( nosim )
  mns <- simulationValues[[1]]
  vrs <- simulationValues[[2]]
	#Means plot
	plotdata <- cumsum(mns)/1:nosim
	meansGraph <- ggplot( data.frame( x = 1:nosim, y = plotdata), aes (x = x, y = y) ) + geom_line() + geom_hline(yintercept = 5) + ggtitle( "Cumulative Means Simulations")

	#Variance plot
  varGraph <- ggplot( data.frame( v = vrs), aes ( x = v ) ) + geom_histogram() + geom_vline(xintercept = 25, aes( colour = "red")) + ggtitle( "Variance Histogram")

	list( meansGraph, varGraph, mns, vrs )
}
```

#Sample mean vs. Theoretical mean

The Central Limit Theorem (CLT) states that sample means follow a normal distribution. Therefore, since we know
the exponential distribution has a mean of $$\frac {1}{\lambda}$$.  The theoretical mean for this simulation is $$\frac {1} {0.2}$$ or 5.
There is a horizontal line at 5 on the simulation graph representing the theoretical mean.  The graph shows, as we
increase the number of simulations, the closer we get to the theoretical mean.

Below is the graph returned by the code above.

```{r, echo=F, fig.width=9, fig.height = 6, fig.align='center'}
graphs <- simulationGraphs()
mnsg <- graphs[[1]]
vrsg <- graphs[[2]]
mnsg
```


#Sample variance vs. Theoretical variance
The theoretical standard deviation is also 1/lambda so the theorectical variance (1/0.2)^2 or 25.  The variance histogram (below) of
the 1000 simulations is centered around 25.




```{r, echo=F, fig.width=9, fig.height = 6, fig.align='center'}
vrsg
```


#Distritbution

The CLT states that the formula

$$\frac{\bar X_n - \mu}{\sigma / \sqrt{n}}$$

is normal.  Using 5 as the theoretical mean and 5 as the theoretical standard deviation, we will run n simulations, subtract 5 and divide by 5/\sqrt{n}.

We will do this for 1,000, 5,000, and 10,000 simulations and plot the results.  Each graph fits a normal distribution.


```{r, echo=F, fig.width=9, fig.height = 6, fig.align='center'}
library(dplyr)

nosims1 <- 1000
nosims2 <- 5000
nosims3 <- 10000

returnedVals <- simulations(nosims1)
mns1 <- returnedVals[[1]]
mns1 <- c(mns1, rep(nosims1, nosims1) )
mns1 <- matrix(mns1, nrow = nosims1, ncol = 2)

returnedVals <- simulations(nosims2)
mns2 <- returnedVals[[1]]
mns2 <- c(mns2, rep(nosims2, nosims2) )
mns2 <- matrix(mns2, nrow = nosims2, ncol = 2)

returnedVals <- simulations(nosims3)
mns3 <- returnedVals[[1]]
mns3 <- c(mns3, rep(nosims3, nosims3) )
mns3 <- matrix(mns3, nrow = nosims3, ncol = 2)

function(x, n) { (x -5) * sqrt(n) / 5}
totalSims <- data.frame( rbind( mns1, mns2, mns3) )
names(totalSims) <- c('means', 'simulations')

cfunc <- function(x, n) { (x -5) * sqrt(n) / 5}
totalSims <- totalSims %>% mutate( a = cfunc(means, simulations))
totalSims$simulations <- factor(totalSims$simulations)

ggplot( totalSims, aes( x = a, fill = simulations) ) + geom_histogram( aes( y = ..density..) ) +  facet_grid(. ~ simulations) ++ ggtitle("Simulations")
```



#Conclusion
The graphs in this report demonstrate the CLT applies to the sample means for the exponental distribution.
